---
title: "Data analysis for plates"
output:
  html_document: default
---

<br>

This notebook predicts the off-target, heterozygous and homozygous samples for two different CRISPR experiments.

<br>

- Select values for global variables:

    - *PLOIDY*: indicates the ploidy of the cell line studied.
    - *EXCLUDED_SAMPLES*: substrings in the sample names that need to be excluded.
    - *EXPERIMENTAL_CONDITION*: triple plasmid CRISPR approach (PPP), plasmid-donor combined with RNPs strategy (PRNP).
    
<br>


```{r}
PLOIDY <- 3

EXCLUDED_SAMPLES <- c("U2OS", "Nup96", "NTC", "Comp")

EXPERIMENTAL_CONDITION <- "PPP"
#EXPERIMENTAL_CONDITION <- "PRNP"
```

- Load libraries and functions.
```{r, warning=F, message=F}
library(dplyr)
source("functions.R")
```

---

### Samples

---

- File paths for each experimental condition.

```{r, echo=F}
if (EXPERIMENTAL_CONDITION=="PPP")
{
  fullGFP.filename <- file.path(getwd(), "data", "plates", paste0("190513_allGFP_HK_Nup93_mEGFP_",EXPERIMENTAL_CONDITION,".csv"))
  fullHDR.filename <- file.path(getwd(), "data", "plates", paste0("190520_HDR_HK_Nup93_mEGFP_",EXPERIMENTAL_CONDITION,".csv"))
}
if (EXPERIMENTAL_CONDITION=="PRNP")
{
  fullGFP.filename <- file.path(getwd(), "data", "plates", paste0("190513_allGFP_HK_Nup93_mEGFP_",EXPERIMENTAL_CONDITION,"-7AS-HiGFP",".csv"))
  fullHDR.filename <- file.path(getwd(), "data", "plates", paste0("190516_HDR_HK_Nup93_mEGFP_",EXPERIMENTAL_CONDITION,"-7AS-HiGFP",".csv"))
}
```

- Read and label data:
```{r}
fullGFP <- read.csv(fullGFP.filename, header=T, stringsAsFactors=F) 
fullHDR <- read.csv(fullHDR.filename, header=T, stringsAsFactors=F) 
fullGFP$Marker <- "GFP"
fullHDR$Marker <- "HDR"
```

- Select relevant columns (**Sample**, **Positives**, **Negatives**, **TargetType**, **Marker**, **T7**, **singleColony**, **imaging**) and bind GFP and HDR data.

```{r, echo=F}
fullGFP <- fullGFP[c("Sample", "Positives", "Negatives", "TargetType", "Marker", "T7", "singleColony", "imaging")]
fullHDR <- fullHDR[c("Sample", "Positives", "Negatives", "TargetType", "Marker", "T7", "singleColony", "imaging")]
dataSAMPLES <- rbind(fullGFP, fullHDR)
rm(fullGFP, fullHDR)
```

- Label the **TargetType** properly:
```{r}
dataSAMPLES[dataSAMPLES$TargetType=="Unknown",]$TargetType <- "TARGET"
dataSAMPLES[dataSAMPLES$TargetType=="Reference",]$TargetType <- "REFERENCE"
```

- Compute **Ratio**, the Confidence Interval 95% (**MinCI95**, **MaxCI95**), relative error (**RelErr**) and theoretical standard deviation (**tSD**):
```{r}
dataSAMPLES$NegativeRate <- computeNegativeRate(dataSAMPLES)
dataSAMPLES <- computeConfidenceInterval(dataSAMPLES)
dataSAMPLES <- computeRelativeError(dataSAMPLES)
dataSAMPLES$tSD <- (dataSAMPLES$Ratio * dataSAMPLES$RelErr) / 1.96
```
> Number of samples = `r nrow(dataSAMPLES)`

- Remove samples with 0:
```{r}
dataSAMPLES <- dataSAMPLES[dataSAMPLES$Positives!=0,]
```

> Number of samples = `r nrow(dataSAMPLES)`

- Remove samples with no values in T7:
```{r}
dataSAMPLES <- dataSAMPLES[!is.na(dataSAMPLES$T7),]
```

> Number of samples = `r nrow(dataSAMPLES)`

- Remove samples with Relative Error > 20 %:
```{r}
dataSAMPLES <- dataSAMPLES[dataSAMPLES$RelErr < 0.2,]
```

> Number of samples = `r nrow(dataSAMPLES)`

- How many samples per TargetType/Marker?
```{r}
knitr::kable(table(dataSAMPLES[c("Marker","TargetType")]))
```

---

### Reference distribution

---

- Read reference data.

```{r, echo=F}
referenceHDR.filename <- file.path(getwd(),  "data", "plates", "reference", "190613_HDR_reference.csv")
refSAMPLES.HDR <- read.csv(referenceHDR.filename, header=T, stringsAsFactors=F) 
referenceGFP.filename <- file.path(getwd(),  "data", "plates", "reference", "190613_allGFP_reference.csv")
refSAMPLES.GFP <- read.csv(referenceGFP.filename, header=T, stringsAsFactors=F) 
refSAMPLES.HDR$Marker <- "GFP"
refSAMPLES.GFP$Marker <- "HDR"
refSAMPLES <- rbind(refSAMPLES.HDR, refSAMPLES.GFP)

refSAMPLES <- convertToLongFormat(refSAMPLES)
```

- Label the **TargetType** of the reference sample.

```{r, echo=F}
refSAMPLES[refSAMPLES$TargetType=="Unknown",]$TargetType <- "TARGET"
refSAMPLES[refSAMPLES$TargetType=="Reference",]$TargetType <- "REFERENCE"
```

- Remove samples with 0 positives:

```{r}
refSAMPLES <- refSAMPLES[refSAMPLES$Positives!=0,]
``` 

- Compute the **Negative Rate** of the reference sample:

```{r}
refSAMPLES$NegativeRate <- computeNegativeRate(refSAMPLES)
```

- Compute the **Ratio** and the **Confidence Interval** of the reference sample:

```{r}
refSAMPLES <- computeConfidenceInterval(refSAMPLES)
```

- Compute **Relative Error**:

```{r}
refSAMPLES <- computeRelativeError(refSAMPLES)
```

- Reference **means** and **SD** for both GFP and HDR dPCR assays:

```{r}
refSAMPLES_GFP_mean <- mean(refSAMPLES[refSAMPLES$Marker=="GFP",]$Ratio)
refSAMPLES_GFP_sd <- sd(refSAMPLES[refSAMPLES$Marker=="GFP",]$Ratio)

refSAMPLES_HDR_mean <- mean(refSAMPLES[refSAMPLES$Marker=="HDR",]$Ratio)
refSAMPLES_HDR_sd <- sd(refSAMPLES[refSAMPLES$Marker=="HDR",]$Ratio)
``` 

---

### Statistical Analysis

---

- Compute **Z-score**: 
```{r}
dataSAMPLES.GFP <- dataSAMPLES[dataSAMPLES$Marker=="GFP",]
dataSAMPLES.HDR <- dataSAMPLES[dataSAMPLES$Marker=="HDR",]

dataSAMPLES.GFP$Zscore <- (dataSAMPLES.GFP$Ratio - refSAMPLES_GFP_mean) / refSAMPLES_GFP_sd
dataSAMPLES.HDR$Zscore <- (dataSAMPLES.HDR$Ratio - refSAMPLES_HDR_mean) / refSAMPLES_HDR_sd
```

- Corrected **p-values** for each sample (by marker) based on the Z-score.

The function `pnorm` provides the integral of the normal curve from `-Inf` to `Z`. Since we are running a two-tailed test, the result should be multiplied by 2. To ensure a negative Z-score, the negative of the absolute value of Z will be computed.

```{r}
dataSAMPLES.GFP$Pvalue <- 2 * pnorm(-abs(dataSAMPLES.GFP$Zscore))
dataSAMPLES.HDR$Pvalue <- 2 * pnorm(-abs(dataSAMPLES.HDR$Zscore))

dataSAMPLES.GFP$correctedPvalue <- p.adjust(dataSAMPLES.GFP$Pvalue, method="fdr")
dataSAMPLES.HDR$correctedPvalue <- p.adjust(dataSAMPLES.HDR$Pvalue, method="fdr")
```

- Selection of good sample candidates based on FDR > 0.1.
```{r}
dataSAMPLES.GFP$GoodSample <- dataSAMPLES.GFP$correctedPvalue >= 0.1
dataSAMPLES.HDR$GoodSample <- dataSAMPLES.HDR$correctedPvalue >= 0.1
```

- Combine GFP and HDR data:
```{r}
dataSAMPLES <- merge (dataSAMPLES.GFP, dataSAMPLES.HDR, by=c("Sample", "TargetType", "T7", "singleColony", "imaging"), suffixes = c(".GFP", ".HDR"))
```

- Keep only the samples with more than 10000 droplets (positives + negatives):
```{r}
dataSAMPLES <- dataSAMPLES %>% mutate (nDroplets_GFP = Positives.GFP+Negatives.GFP) %>%
                             mutate (nDroplets_HDR = Positives.HDR+Negatives.HDR) %>%
                             filter (nDroplets_GFP > 10000 & nDroplets_HDR > 10000)
```

####  Predict outcome

  - *AGREEMENT* if Zeta <= 2
  - *WARNING* if Zeta $\in$ (2,3]
  - *DISAGREEMENT* if Zeta > 3
  
```{r}
dataSAMPLES$Zeta <- abs(dataSAMPLES$Ratio.GFP - dataSAMPLES$Ratio.HDR) / 
                   sqrt(dataSAMPLES$tSD.GFP^2 + dataSAMPLES$tSD.HDR^2)

dataSAMPLES <- mutate(dataSAMPLES, prediction = ifelse(Zeta <= 2, "AGREEMENT", 
                                              ifelse ((Zeta > 2 & Zeta <= 3), "WARNING", "DISAGREEMENT")))
```

- Potential off-targets (**singleColony** is 1 when the sample is accepted):
```{r}
offtargetSamples <- dataSAMPLES %>% filter(prediction == "DISAGREEMENT" & (Ratio.GFP > Ratio.HDR) & (dataSAMPLES$singleColony == 1))
``` 

- Potential heterozygous:
```{r}
heteroSamples <- dataSAMPLES %>% filter(prediction == "AGREEMENT" & GoodSample.GFP == FALSE & GoodSample.HDR == FALSE & dataSAMPLES$singleColony == 1)
``` 

- Potential homozygous:
```{r}
homoSamples <- dataSAMPLES %>% filter(prediction == "AGREEMENT" & GoodSample.GFP == TRUE & GoodSample.HDR == TRUE & dataSAMPLES$singleColony == 1)
``` 

- Unclassified (excluded):
```{r}
classifiedSamples <- rbind (offtargetSamples, heteroSamples, homoSamples) 
excludedSamples <- dataSAMPLES %>% filter(!Sample %in% classifiedSamples$Sample) %>% select(Sample, Ratio.GFP, Ratio.HDR, GoodSample.GFP, GoodSample.HDR, prediction, T7, singleColony)
``` 

> `r nrow(classifiedSamples)` samples classified and `r nrow(excludedSamples)` excluded.


#### Pie chart of off-targets, heterozygous and homozygous clones


```{r}
pie_values <- c(nrow(offtargetSamples), nrow(heteroSamples), nrow(homoSamples))
sampleSize <- sum(pie_values)

pie_labels <- paste(c("off-target", "heterozygous", "homozygous"), " = " , pie_values)
pie(pie_values, labels=pie_labels, main=paste("Exp. Condition:",EXPERIMENTAL_CONDITION, "\n", "Sample number =", sampleSize))
```
